A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2021-11-02 paper/elsarticle/elsarticle-template.tex
paper

\begin{abstract}
Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to such instructions in paper form or on a statically situated screen. \textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for any to hold a handheld device such as a tablet or for each to wear some form of head-mounted display. In carrying out spatial AR, however, care must be taken to appropriately warp an image to be projected in a manner that it, when projected, appear free of distortions to the viewer. For planar scene geometry (such as a floor, wall, or table), a manual process called keystone correction can be used to carry out an appropriate corrective image warp, a process that is cumbersome and time consuming. Another drawback of conventional projector-based spatial AR is that it is capable of augmenting only the portion of the scene within the projector's own static field of view, thereby hindering its applicability to use cases calling for augmentation of wide areas such as a factory floorspace.

We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically as a function of the relative geometry of the projector and scene plane, using a projector equipped with a steerable mirror to direct the projection across varying target locations and a camera facing the scene plane in support of calibrating the system. Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways. First, (i) the horizontal and vertical axes of the desired augmentations is carried out in accordance with the horizontal and vertical image axes of the camera, thereby making setting those axes as simple as appropriately rotating the camera relative to the scene. Second, (ii) the desired dimensions of the projected augmentations are specified in metric terms, thereby allowing for consistent scaling across all target locations. %Our system produces such augmentations accurately, with compelling time savings relative manual keystone correction.
\end{abstract}

@@

Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene~\cite{schlund2018moglichkeiten,uva2018evaluating,masood2019augmented,gattullo2019towards,aschenbrenner2019comparing,mayrhofer2019one,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to such instructions in paper form or on a statically situated screen. Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed typically using some form of handheld or head-mounted display. Handheld devices such as tablets do not enable the user to work hands-free while using the device, or force the user to wear additional gear. Reliance on head-mounted displays has two adverse consequences of its own: a head-mounted display must be worn by each individual wishing to partake in the augmentation, and such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive. A HoloLens 2 (Microsoft Corporation, USA), for instance, is such a head-mounted display; at a weight of 566~grams \cite{hololens}, it places an additional ergonomic load on the worker, in particular if worn over a longer duration or in bent positions.

In turn, a form of augmented reality referred to as \textit{spatial} AR\cite{bimber2019spatial} is carried out not by embedding the augmentation in an image of the scene as with a head-mounted or handheld display, but by projection to the scene itself, thus eliminating the aforementioned problems. Spatial AR thus turns surfaces such as the product being manufactured, a desk, the floor, a wall, or even the ceiling (considering overhead work) into displays. Information provision---of the work plan or of quality-related information---can be carried out by projection to the given workspace, while directly referring to exact positions within that workspace. Moreover, the fact that this information can be viewed by multiple users simultaneously serves to promote collaboration \cite{aschenbrenner2019comparing}.

\begin{figure}
    \centering
    \subfloat[\centering Original image.]{{\includegraphics[height=3.4cm]{images/projim.png}}}
    \qquad
    \subfloat[\centering Projecting the original image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
    \caption{Projection to the scene plane from the viewpoint of the recovered projector, facing a given target location. (a) Original image. (b) Projecting the original image from the viewpoint of the recovered projector (bottom left, black) to the scene plane introduces projective distortions if the projector is not fronto-parallel with the scene plane, as is the case here.}
    \label{fig:proj}
\end{figure}

While interest in spatial AR in industrial settings has been growing, its deployment in industrial settings remains well behind that of wearable AR systems such as handhelds or head-mounted displays. One of the main shortcomings of typical projector-based spatial AR relative to AR relying on wearables is the applicability of spatial AR only to augmenting scenes within the projector's own static field of view. Here, dynamic projection systems---using mirror heads to deflect the projection across wider areas---has the potential to bridge the gap to wider industrial penetration. Yet the increase in the area to be covered comes with noteworthy challenges in terms of ensuring that the images to be projected appear free of distortions to the users. Considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions (cf.\ Figure~\ref{fig:proj}). Such distortions can be corrected for by carrying out a cumbersome manual process referred to as keystone correction---involving warping an image by perturbing its corners until the desired effect is produced when projected to the scene---to appropriately warp the image to be projected, often using software bundled with the projector. Automating this manual, time-consuming, and error-prone process significantly reduces the time needed for the industrial engineering of a spatial AR application and therefore has potential to promote adoption by industry. Especially manufacturing processes of large products that remain at a static mounting position during the assembly process---i.e., industrial site assembly---offer a significant application potential for the use of spatial AR as part of worker assistance systems \cite{mayrhofer2019one}.

\begin{figure}
    \centering
    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projecting our warped image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
    \caption{Projection to the scene plane from the viewpoint of the recovered projector for a given target location. (a) Original image warped using our approach. (b) After warping the original image according to our approach for the given target location, the image is projected from the viewpoint of the recovered projector (bottom left, black) with the effect of projecting the original image from the viewpoint of the virtual projector (center, gray). This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale. Note that black pixels will in effect be ignored by the projector.}
    \label{fig:warp}
\end{figure}

Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction. Most importantly, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face directly towards the scene plane (a camera we shall in what follows refer to as `downwards facing', for brevity). This facilitates placement of augmentations, since establishing the principal axes of the augmentation thus reduces to placing the downwards-facing camera in a manner such that the $X$- and $Y$-axes of the image align with the intended $X$- and $Y$-axes of the augmentation. We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}). We place this virtual projector to (i) face directly downwards to the target location in the scene plane and (ii) rotate its axes in accordance with those of the camera. Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane. Finally, our system set up to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.

@@

An early spatial AR system explicitly using a projector mounted with a steerable mirror is the IBM Everywhere Displays prototype of Pinharez \cite{pinhanez2001everywhere}. For the purposes of the prototype, a projector mounted with a steerable mirror was set up with a camera to demonstrate a variety of use cases, including collaborative assembly encompassing the projection of assembly instructions and an interactive projected user interface \cite{kjeldsen2002interacting,pinhanez2003applications}. The authors, however, carry out keystone correction manually, by interactively adjusting for the position and orientation of a virtual scene plane and a scaling factor and computing a 2D homography accordingly. %Rather than rely on a steerable mirror to support spatial AR to multiple locations using a single projector, some works mount the projector and a camera on a rigid rig and subject the rig to motion \cite{ehnes2004projected,borkowski2004spatial,butz2006applying}. Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to rigid motion.

More generally, keystone correction for planar scenes can be said to reduce to computing a 2D homography \cite{Hartley2004}, an invertible transformation that preserves colinearity and is thus fittingly alternatively referred to as a `colineation'; this is the case whether computing the homography relies on manual interaction as in the case of Pinharez, or is obtained free of manual interaction as in ours. The intuition for why it is that a transformation that preserves colinearity---i.e., maps lines to lines---can serve to model an appropriate corrective image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity'). To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity) calls for a transformation that maps lines to lines.

One way to compute a homography is by identifying at least four correspondences between pixel positions in two images of a planar surface \cite{Hartley2004}. The keystone correction approach of Sukthankar \textit{et al.} \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane. Additional sensors can be used to inform the computation of a homography; Raskar and Beardsley \cite{raskar2001self} use a tilt sensor to recover the projector's gravity vector, which they use to carry out a rotational correction in projecting to a wall. Still another way to compute a homography is as a function of a plane and a pair of pinhole cameras; the effect produced by the homography is one of projecting an image to the plane from the viewpoint of the one camera, and acquiring the projected image from the viewpoint of the other. Such a homography is said to be `induced' by a plane \cite{Hartley2004}. We exploit the fact that cameras and projectors alike can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp, taking into account both desired in-plane orientation and metric scale.

@@

Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} can be achieved by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view. This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene surface; modeling the geometry that relates projector and scene plane thus enables one to reason about the problem of correcting for distortions in terms of geometry. To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downwards to the scene plane and includes recovery of the scene plane as a convenient side effect. Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that warps the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene. These two points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.

@@

Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which a projecting point~$\mathbf{X}$ must lie. Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length~$f$ and principal point~$\mathbf{p}_0 = (x_0, y_0)^\top \in \mathbb{R}^2$ \cite{Hartley2004}, and (ii)~the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close,weng1992camera}; the calibration matrix~$\mathtt{K}$ is then
\begin{equation}
\mathtt{K} = \begin{bmatrix}
f & 0 & x_0 \\
0 & f & y_0 \\
0 & 0 & 1
\end{bmatrix},
\end{equation}
where $f$ and $(x_0, y_0)^\top$ are both expressed in units of pixels. The calibration matrix~$\mathtt{K}$ models a so-called pinhole camera; such a model is applicable to an image acquired using a real-world camera assuming the image has been corrected for lens distortion effects, by having applied the appropriate lens distortion model coefficients.

@@

Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; if proceeding accordingly, the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the \textit{projector} and the corresponding points in the scene. What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the downwards-facing camera (used to obtain the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (used for obtaining the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$, for each projector calibration image).

@@

\begin{figure}[ht]
    \centering
    \subfloat[\centering Projector calibration image (detections overlain).]{{\includegraphics[height=3cm]{images/detections.png} }}
    \qquad
    \subfloat[\centering 3D positions for projector calibration (red, in scene plane).]{{\includegraphics[height=3cm]{images/2d3d.png}}}
    \caption{Obtaining the 3D positions of the 2D-3D correspondences for projector calibration. (a) For each target location, a projector calibration image is acquired from the viewpoint of the downwards-facing camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby. Automatically detected circle pattern center points and chessboard corners overlain for illustration. (b) Given a projector calibration image, the scene plane (gray) is recovered in the coordinate frame of the camera (black, up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used for calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown). The downwards-facing camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
    \label{fig:2d}
\end{figure}

@@

It is, strictly speaking, in the sense that we set the $X$- and $Y$-axes of the virtual projector to the $X$- and $Y$-axes of the virtual camera that we place the horizontal and vertical axes of the desired augmentations in accordance with the horizontal and vertical image axes of the camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed. Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector. A steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, thereby facilitating placement of target locations.

@@

\section{Evaluation}
We evaluate our approach by using a 960 pixel~$\times$~600 pixel image of the TU Wien logo (cf.\ Figure \ref{fig:warp}) to augment 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria. We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace (i.e., the yellow and red lines in Figure~\ref{fig:eval}(b)), absent of projective distortions, and with the same metric dimensions of 50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}. All experiments were carried out by the same technician, experienced in both approaches.

The hardware setup employed in the evaluation (cf.\ Figure \ref{fig:eval}(a)) comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used in our experiments to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information,Rupprecht2021}. The steerable mirror system was bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset. In addition, we used a downwards-facing Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view. The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m. At this height, the pixel size (i.e., length or width of a pixel if projected to the scene plane) of the camera was ca.\ 3.5 mm.

>>>>

\begin{abstract}
Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to such instructions in paper form or on a statically situated screen. \textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for any to hold a handheld device such as a tablet or for each to wear some form of head-mounted display. In carrying out spatial AR, however, care must be taken to appropriately warp the image to be projected such that it, when projected, appear free of projective distortions to the viewer. For planar scene geometry (such as a floor, wall, or table), a manual process called keystone correction can be used to carry out an appropriate corrective image warp, a process that can be cumbersome. Another drawback of conventional spatial AR relying only on a single projector is that it is capable of augmenting only the portion of the scene within the projector's static field of view, thereby hindering its applicability to use cases calling for augmentation of wide areas such as a factory floorspace.

We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically as a function of the relative geometry of the projector and scene plane, using a projector equipped with a steerable mirror to direct the projection across varying target locations and a camera facing the scene plane in support of calibrating the system. Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways. First, (i) the horizontal and vertical axes of the desired augmentations is carried out in accordance with the horizontal and vertical image axes of the camera, thereby making setting those axes as simple as appropriately rotating the camera relative to the scene plane, and thereby providing a consistent axial reference frame across all target locations. Second, (ii) the desired dimensions of the projected augmentations are specified in metric terms, thereby providing for consistent scaling, likewise across all target locations. %Our system produces such augmentations accurately, with compelling time savings relative manual keystone correction.
\end{abstract}

@@

Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene~\cite{schlund2018moglichkeiten,uva2018evaluating,masood2019augmented,gattullo2019towards,aschenbrenner2019comparing,mayrhofer2019one,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to such instructions in paper form or on a statically situated screen. Typically, AR works by embedding the augmentation in imagery of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed typically using some form of handheld or head-mounted display. Handheld devices such as tablets do not enable the user to work hands-free while using the device, or force the user to wear additional gear. Reliance on head-mounted displays has two adverse consequences of its own: a head-mounted display must be worn by each individual wishing to partake in the augmentation, and such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive. A HoloLens 2 (Microsoft Corporation, USA), for instance, is such a head-mounted display; at a weight of 566~grams \cite{hololens}, it places an additional ergonomic load on the worker, in particular if worn over a longer duration or in bent positions.

In turn, a form of augmented reality referred to as \textit{spatial} AR\cite{bimber2019spatial} is carried out not by embedding the augmentation in an image of the scene as with a head-mounted or handheld display, but by projection to the scene itself, thus eliminating the aforementioned problems. Spatial AR thus turns surfaces such as the object being manufactured, a desk, the floor, a wall, or even the ceiling (considering overhead work) into displays. Information provision---of the work plan or of quality-related information---can be carried out by projection to the given workspace, while directly referring to exact positions within that workspace. Moreover, the fact that this information can be viewed by multiple users simultaneously serves to facilitate collaboration \cite{aschenbrenner2019comparing}.

\begin{figure}
    \centering
    \subfloat[\centering Original image.]{{\includegraphics[height=3.4cm]{images/projim.png}}}
    \qquad
    \subfloat[\centering Projection to target location in scene plane (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
    \caption{Projection to a given target location in the scene plane (e.g., the floor, a wall, or table). (a) Original image. (b) Projecting the original image from the viewpoint of the projector (bottom left, black) to the scene plane introduces projective distortions if the projector is not fronto-parallel with the scene plane, as is the case here.}
    \label{fig:proj}
\end{figure}

While interest in spatial AR in industrial settings has been growing, its deployment in industrial settings remains well behind that of wearable AR systems such as handhelds or head-mounted displays. One of the main shortcomings of typical projector-based spatial AR relative to AR relying on wearables is the ability to project augmentations only within the projector's own static field of view. Here, dynamic projection systems---using steerable mirror heads to direct the projection across wider areas---has the potential to bridge the gap to wider industrial penetration. Yet the increase in the area to be covered comes with noteworthy challenges in terms of preparing the augmentations, in particular as the number of target locations in the scene becomes larger than a small handful. Considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions (cf.\ Figure~\ref{fig:proj}). Such distortions can be corrected for by carrying out a cumbersome manual process referred to as keystone correction---involving warping an image by perturbing its corners until the desired effect is produced when projected to the scene---to appropriately warp the image to be projected, often using software bundled with the projector. Automating this manual, time-consuming, and error-prone process significantly reduces the time needed for the industrial engineering of a spatial AR application and therefore has potential to promote adoption by industry. Especially manufacturing processes of large products that remain at a static mounting position during the assembly process---i.e., industrial site assembly---offer a significant application potential for the use of spatial AR as part of worker assistance systems \cite{mayrhofer2019one}.

\begin{figure}
    \centering
    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projection to target location in scene plane (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
    \caption{Projection to a given target location in the scene plane. (a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped with respect to the given target location using our approach. (b) Projecting the warped image from the viewpoint of the projector (bottom left, black) has the effect of projecting the \textit{original} image from the viewpoint of the virtual projector (center, gray). This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale. Note that black pixels will in effect be ignored by the projector.}
    \label{fig:warp}
\end{figure}

Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction. Notably, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face directly towards the scene plane (a camera we shall in what follows refer to as `downwards facing', for brevity). This facilitates placement of augmentations, since establishing the principal axes of the augmentation thus reduces to appropriately rotating the downwards-facing camera relative to the scene plane. We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}). We place this virtual projector to (i) face directly downwards to the target location in the scene plane and (ii) rotate its axes in accordance with those of the camera. Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane. Finally, our system is set up to handle a projector equipped with a steerable mirror in a manner that does not call for explicitly modeling the action of the steerable mirror on the projector, thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.

@@

An early spatial AR system explicitly using a projector mounted with a steerable mirror is the IBM Everywhere Displays prototype of Pinharez \cite{pinhanez2001everywhere}. For the purposes of the prototype, a projector mounted with a steerable mirror was set up with a camera to demonstrate a variety of use cases, including collaborative assembly encompassing the projection of assembly instructions and an interactive projected user interface \cite{kjeldsen2002interacting,pinhanez2003applications}. The authors, however, carry out keystone correction manually, by interactively adjusting for the position and orientation of a virtual scene plane and a scaling factor and computing a corresponding 2D homography accordingly. %Rather than rely on a steerable mirror to support spatial AR to multiple locations using a single projector, some works mount the projector and a camera on a rigid rig and subject the rig to motion \cite{ehnes2004projected,borkowski2004spatial,butz2006applying}. Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to rigid motion.

More generally, keystone correction for planar scenes can be said to reduce to computing a 2D homography \cite{Hartley2004}, an invertible transformation that preserves colinearity and is thus fittingly alternatively referred to as a `colineation'; this is the case whether computing the homography relies on manual interaction as in the case of Pinharez, or is obtained free of manual interaction as in ours. The intuition for why it is that a transformation that preserves colinearity---i.e., maps lines to lines---can serve to model an appropriate corrective image warp can be drawn from considering a planar chessboard pattern: looking at a chessboard from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points, yet remain lines; looking at the same chessboard frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity'). To warp an image of the chessboard acquired at one angle such that the lines of the chessboard appear as if the chessboard were acquired at another angle accordingly calls for a transformation that maps lines to lines; it therefore follows that there exists a homography that models that transformation.

One way to compute a homography is by identifying at least four correspondences between pixel positions in two images of a planar surface \cite{Hartley2004}. The keystone correction approach of Sukthankar \textit{et al.} \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane. Additional sensors can be used to inform the computation of a homography; Raskar and Beardsley \cite{raskar2001self} use a tilt sensor to recover the projector's gravity vector, which they use to carry out a rotational correction in projecting to a wall. Still another way to compute a homography is as a function of a plane and a pair of pinhole cameras; the effect produced by the homography is one of projecting an image to the plane from the viewpoint of the one camera, and acquiring the projected image from the viewpoint of the other. Such a homography is said to be `induced' by a plane \cite{Hartley2004}. We exploit the fact that cameras and projectors alike can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp, yet additionally taking into account both desired in-plane orientation and metric scale.

@@

Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro}---and exemplified in Figure~\ref{fig:proj}---can be achieved by means of an appropriate plane-induced homography \cite{Hartley2004}, computed by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view. This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene plane; modeling the geometry that relates projector and scene plane thus enables one to reason about the problem of correcting for distortions in terms of geometry. To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downwards to the scene plane and includes recovery of the scene plane as a convenient side effect. Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that warps the image to be projected in a manner that it appear undistorted to the viewer, placed in alignment with the axes of a proxy image of the scene, and at the desired metric scale. These two points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.

@@

Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which a projecting point~$\mathbf{X}$ must lie. Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length~$f$ and principal point~$\mathbf{p}_0 = (x_0, y_0)^\top \in \mathbb{R}^2$ \cite{Hartley2004}, and (ii)~the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close,weng1992camera}; the calibration matrix~$\mathtt{K}$ is then
\begin{equation}
\mathtt{K} = \begin{bmatrix}
f & 0 & x_0 \\
0 & f & y_0 \\
0 & 0 & 1
\end{bmatrix},
\end{equation}
where $f$, $x_0$, and $y_0$ are each expressed in units of pixels. The calibration matrix~$\mathtt{K}$ models a so-called pinhole camera \cite{Hartley2004}; such a model is applicable to an image acquired using a real-world camera assuming the image has in a prior step been corrected for lens distortion effects. %, by having applied the appropriate lens distortion model coefficients.

@@

Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera then concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the \textit{projector} and the corresponding points in the scene. What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the downwards-facing camera (giving the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (giving the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$, for each projector calibration image).

@@

\begin{figure}[ht]
    \centering
    \subfloat[\centering Projector calibration image (detections overlain).]{{\includegraphics[height=3cm]{images/detections.png} }}
    \qquad
    \subfloat[\centering 3D positions for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3cm]{images/2d3d.png}}}
    \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration. (a) For each target location, a projector calibration image is acquired from the viewpoint of the downwards-facing camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby. Circle pattern center points and chessboard corners detected automatically (overlain for illustration). (b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the camera (black) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown). The downwards-facing camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
    \label{fig:2d}
\end{figure}

@@

It is, strictly speaking, in the sense that we set the $X$- and $Y$-axes of the virtual projector to the $X$- and $Y$-axes of the virtual camera that we place the horizontal and vertical axes of the desired augmentations in accordance with the horizontal and vertical image axes of the camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed. Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector. A steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, thereby facilitating placement.

@@

\section{Evaluation}
We evaluate our approach by using a 960 pixel~$\times$~600 pixel image of the TU Wien logo (cf.\ Figure \ref{fig:warp}(a)) to augment 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria. We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace (i.e., the yellow and red lines in Figure~\ref{fig:2d}(b)), absent of projective distortions, and with the same metric dimensions of 50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}. All experiments were carried out by the same technician, experienced in both approaches.

The hardware setup employed in the evaluation (cf.\ Figure \ref{fig:eval}(a)) comprised a Panasonic PT-RZ660BE projector (Panasonic Corporation, Japan) with a steerable mirror system---used in our experiments to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute (Dynamic Projection Institute, Austria) \cite{rupprecht2020information,Rupprecht2021}. The steerable mirror system was bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset. In addition, we used a downwards-facing Stereolabs Zed 2 stereo camera (Stereolabs Inc., France), yet relied only on the left view. The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m. At this height, the pixel size (i.e., length or width of a pixel if projected to the scene plane) of the camera was ca.\ 3.5 mm.
